### 2022-04-06 stats/supplements_webcamcoding.Rmd
create saves folder, add webcam coding script

---
title: "supplements_webcamcoding"
author: "Julia Prein"
date: '2022-04-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(brms)
library(ggridges)

options(scipen = 999)
theme_set(theme_classic())
```

```{r load_data}
webcam_data <- readRDS(file = "../data/gafo-testtrials.rds") %>% 
  filter(studytype == "vali" & datacollection == "remote - unsupervised" & studyversion == "box" & webcam == TRUE & ageInYears == 3) %>% 
  group_by(subjID) %>% 
  # determine order of factors
  mutate(
    targetPosition = factor(targetPosition, levels = c("box1", "box2", "box3", "box4", "box5")), 
    proportionCorrectBox = mean(correctBox, na.rm = TRUE),
  )
```

### VISUAL EXPLORATION

```{r webcam_exploring}
### nrRepeatedTestquestion
webcam_data %>% 
  group_by(subjID) %>% 
  mutate(repeatedTestquestion = mean(nrRepeatedTestquestion, na.rm = TRUE)) %>% 
  ggplot(., aes(x = repeatedTestquestion, y = proportionCorrectBox)) + 
  geom_point(col = "#006c66", size = 0.7) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.01) + 
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))

### nrHintsEyes
webcam_data %>% 
  group_by(subjID) %>% 
  mutate(hintsEyes = mean(nrHintsEyes, na.rm = TRUE)) %>% 
  ggplot(., aes(x = hintsEyes, y = proportionCorrectBox)) + 
  geom_point(col = "#006c66", size = 0.7) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.01) + 
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))

### nrVerificationChoice
webcam_data %>% 
  group_by(subjID) %>% 
  mutate(verificationChoice = mean(nrVerificationChoice, na.rm = TRUE)) %>% 
  ggplot(., aes(x = verificationChoice, y = proportionCorrectBox)) + 
  geom_point(col = "#006c66", size = 0.7) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.01) + 
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))

### nrParentSpeech
webcam_data %>% 
  group_by(subjID) %>% 
  mutate(parentSpeech = mean(nrParentSpeech, na.rm = TRUE)) %>% 
  ggplot(., aes(x = parentSpeech, y = proportionCorrectBox)) + 
  geom_point(col = "#006c66", size = 0.7) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.15, col = "darkgrey") +
  stat_cor(r.accuracy = 0.01, p.accuracy = 0.01) + 
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))
```

### MODELS ON TRIAL LEVEL

```{r webcam_perTrial_datapreparation}
# prepare data
mData <- webcam_data %>% 
  ungroup() %>% 
  mutate(
    correct = as.numeric(correctBox),

    # centered around minimum
    age = ageInMonths - min(ageInMonths),

    # centered around mean
    trial = scale(trialNr, center = TRUE, scale = FALSE), 
    repeatedTestquestion = scale(nrRepeatedTestquestion, center = TRUE, scale = FALSE),
    verificationChoice = scale(nrVerificationChoice, center = TRUE, scale = FALSE),
    hintsEyes = scale(nrHintsEyes, center = TRUE, scale = FALSE),
    nrParentSpeech = replace_na(nrParentSpeech, 0), 
    parentSpeech = scale(nrParentSpeech, center = TRUE, scale = FALSE),

    # calculating index
    parentalInterference = -(repeatedTestquestion) + -(verificationChoice) + hintsEyes,

    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    pi = scale(parentalInterference, scale = TRUE, center = TRUE), 
  ) %>% 
  select(trial, subjID, correct, targetPosition, age, parentalInterference, pi, repeatedTestquestion, verificationChoice, hintsEyes, parentSpeech)

```

```{r webcam_perTrial_models}
# originally: 
# brm(correct ~ age + (trial + targetPosition | subjID) + (age | targetPosition)
# age + person effect + trial effect
# simplified the model, since not a lot of data / info to estimate all the parameters. 
# therefore, remove trial & targetposition

mPerTrialNull <- brm(correct ~ age + (1 | subjID) + (1 | targetPosition), 
                              data   = mData, 
                              family = bernoulli(link = "logit"), 
                              warmup = 1000, 
                              iter   = 3000, 
                              chains = 4, 
                              inits  = "random",
                              cores  = 4)

# we decided to build an index since we're rather interested in the additive effect of the parameters; general parental inference
# we don't want to know what happens if we e.g. set repeatedQuestion & verfication to zero how does hintsEyes influence correct

# index: repeated testquestion & verification choice tend to lean into negative effect, 
# hints eyes rather positive. therefore -, - + for summing up
mPerTrialPI <- update(mPerTrialNull,
                      newdata = mData,
                      formula = correct ~ age + pi + (1 | subjID) + (1 | targetPosition), 
                      family = bernoulli(link = "logit"),
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerTrialRepeatedTestquestion <- update(mPerTrialNull,
                      newdata = mData,
                      formula = correct ~ age + repeatedTestquestion + (1 | subjID) + (1 | targetPosition), 
                      family = bernoulli(link = "logit"),
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerTrialVerificationChoice <- update(mPerTrialNull,
                      newdata = mData,
                      formula = correct ~ age + verificationChoice + (1 | subjID) + (1 | targetPosition), 
                      family = bernoulli(link = "logit"),
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerTrialHintsEyes <- update(mPerTrialNull,
                      newdata = mData,
                      formula = correct ~ age + hintsEyes + (1 | subjID) + (1 | targetPosition), 
                      family = bernoulli(link = "logit"),
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerTrialParentSpeech <- update(mPerTrialNull,
                      newdata = mData,
                      formula = correct ~ age + parentSpeech + (1 | subjID) + (1 | targetPosition), 
                      family = bernoulli(link = "logit"),
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)
```

```{r webcam_perTrial_comparisons}
mPerTrialNull <- add_criterion(mPerTrialNull, "waic")
mPerTrialPI <- add_criterion(mPerTrialPI, "waic")
mPerTrialRepeatedTestquestion <- add_criterion(mPerTrialRepeatedTestquestion, "waic")
mPerTrialVerificationChoice <- add_criterion(mPerTrialVerificationChoice, "waic")
mPerTrialHintsEyes <- add_criterion(mPerTrialHintsEyes, "waic")
mPerTrialParentSpeech <- add_criterion(mPerTrialParentSpeech, "waic")

mPerTrialComparison <- loo_compare(mPerTrialNull, mPerTrialPI, mPerTrialRepeatedTestquestion, mPerTrialVerificationChoice, mPerTrialHintsEyes, mPerTrialParentSpeech, criterion = "waic")

mPerTrialComparison <- as_tibble(mPerTrialComparison, rownames = "model")
write.csv(mPerTrialComparison, "../saves/supplements-webcamcoding-mPerTrialComparison-WAIC.csv", quote = F, row.names = F)
saveRDS(mPerTrialComparison, "../saves/supplements-webcamcoding-mPerTrialComparison-WAIC.rds")

mPerTrialComparisonWeights <- as_tibble(model_weights(mPerTrialNull, mPerTrialPI, mPerTrialRepeatedTestquestion, mPerTrialVerificationChoice, mPerTrialHintsEyes, mPerTrialParentSpeech, criterion = "waic"), rownames = "model")
write.csv(mPerTrialComparisonWeights, "../saves/supplements-webcamcoding-mPerTrialComparison-weights.csv", quote = F, row.names = F)
saveRDS(mPerTrialComparisonWeights, "../saves/supplements-webcamcoding-mPerTrialComparison-weights.rds")
```

```{r webcam_perTrial_plot}
postSamplesPerTrial <- bind_rows(
  # posterior_samples(mPerTrialNull, pars = "age") %>% mutate(model = "mPerTrialNull - age") %>% rename(estimate = b_age),
  posterior_samples(mPerTrialPI, pars = "pi") %>% mutate(model = "mPerTrialPI") %>% rename(estimate = b_pi), 
  posterior_samples(mPerTrialRepeatedTestquestion, pars = "repeatedTestquestion") %>% mutate(model = "mPerTrialRepeatedTestquestion") %>% rename(estimate = b_repeatedTestquestion), 
  posterior_samples(mPerTrialVerificationChoice, pars = "verificationChoice") %>% mutate(model = "mPerTrialVerificationChoice") %>% rename(estimate = b_verificationChoice), 
  posterior_samples(mPerTrialHintsEyes, pars = "hintsEyes") %>% mutate(model = "mPerTrialHintsEyes") %>% rename(estimate = b_hintsEyes), 
  posterior_samples(mPerTrialParentSpeech, pars = "parentSpeech") %>% mutate(model = "mPerTrialParentSpeech") %>% rename(estimate = b_parentSpeech), 
)

postSamplesPerTrial <- postSamplesPerTrial %>% 
  mutate(
        model = factor(model, levels = c("mPerTrialParentSpeech", "mPerTrialHintsEyes", "mPerTrialVerificationChoice", "mPerTrialRepeatedTestquestion", "mPerTrialPI")),
  )

ggplot(postSamplesPerTrial, aes(x = estimate, y = model, fill = factor(stat(quantile)))) +
  geom_vline(xintercept = 0, lty = 2, alpha = .75) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
  ) +	
  scale_fill_manual(values = c("firebrick", "NA", "firebrick"), guide = "none") +
  scale_x_continuous(limits = c(-2, 2), breaks = seq(-2, 2, by = 0.5))

ggsave("../figures/supplements-webcamcoding-mPerTrialComparison-posteriorsample.png")
```

### MODELS ON SUBJECT LEVEL

```{r webcam_perSubject_datapreparation}
mDataSubj <- webcam_data %>% 
  group_by(subjID) %>% 
  mutate(
    trials = n(),
    correctTrials = sum(correctBox), 
    repeatedTestquestion = sum(nrRepeatedTestquestion), 
    verificationChoice = sum(nrVerificationChoice), 
    hintsEyes = sum(nrHintsEyes), 
    parentalInterference = -(repeatedTestquestion) + verificationChoice + -(hintsEyes),
    nrParentSpeech = replace_na(nrParentSpeech, 0), 
    parentSpeech = sum(nrParentSpeech), 
  ) %>% 
  slice(1) %>% 
  # for scaling we need ungrouped data!
  ungroup() %>% 
  mutate(
    age = ageInMonths - mean(ageInMonths), 
    pi = scale(parentalInterference, scale = TRUE, center = TRUE), 
  ) %>% 
  select(subjID, correctTrials, trials, age, repeatedTestquestion, verificationChoice, hintsEyes, parentSpeech, pi)
```

```{r webcam_perSubject_models}
mPerSubjectNull <- brm(correctTrials |  trials(trials) ~ age, 
                              data   = mDataSubj, 
                              family = binomial(link = "logit"), 
                              warmup = 1000, 
                              iter   = 3000, 
                              chains = 4, 
                              inits  = "random",
                              cores  = 4)

mPerSubjectPI <- update(mPerSubjectNull,
                      newdata = mDataSubj,
                      formula = correctTrials |  trials(trials) ~ age + pi, 
                      family = binomial(link = "logit"), 
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerSubjectRepeatedTestquestion <- update(mPerSubjectNull,
                      newdata = mDataSubj,
                      formula = correctTrials |  trials(trials) ~ age + repeatedTestquestion, 
                      family = binomial(link = "logit"), 
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerSubjectVerificationChoice <- update(mPerSubjectNull,
                      newdata = mDataSubj,
                      formula = correctTrials |  trials(trials) ~ age + verificationChoice, 
                      family = binomial(link = "logit"), 
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerSubjectHintsEyes <- update(mPerSubjectNull,
                      newdata = mDataSubj,
                      formula = correctTrials |  trials(trials) ~ age + hintsEyes, 
                      family = binomial(link = "logit"), 
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)

mPerSubjectParentSpeech <- update(mPerSubjectNull,
                      newdata = mDataSubj,
                      formula = correctTrials |  trials(trials) ~ age + parentSpeech, 
                      family = binomial(link = "logit"), 
                      warmup = 1000, 
                      iter   = 3000, 
                      chains = 4, 
                      inits  = "random",
                      cores  = 4)
```

```{r webcam_perTrial_comparisons}
mPerSubjectNull <- add_criterion(mPerSubjectNull, "waic")
mPerSubjectPI <- add_criterion(mPerSubjectPI, "waic")
mPerSubjectRepeatedTestquestion <- add_criterion(mPerSubjectRepeatedTestquestion, "waic")
mPerSubjectVerificationChoice <- add_criterion(mPerSubjectVerificationChoice, "waic")
mPerSubjectHintsEyes <- add_criterion(mPerSubjectHintsEyes, "waic")
mPerSubjectParentSpeech <- add_criterion(mPerSubjectParentSpeech, "waic")

mPerSubjectComparison <- loo_compare(mPerSubjectNull, mPerSubjectPI, mPerSubjectRepeatedTestquestion, mPerSubjectVerificationChoice, mPerSubjectHintsEyes, mPerSubjectParentSpeech, criterion = "waic")

mPerSubjectComparison <- as_tibble(mPerSubjectComparison, rownames = "model")
write.csv(mPerSubjectComparison, "../saves/supplements-webcamcoding-mPerSubjectComparison-WAIC.csv", quote = F, row.names = F)
saveRDS(mPerSubjectComparison, "../saves/supplements-webcamcoding-mPerSubjectComparison-WAIC.rds")

mPerSubjectComparisonWeights <- as_tibble(model_weights(mPerSubjectNull, mPerSubjectPI, mPerSubjectRepeatedTestquestion, mPerSubjectVerificationChoice, mPerSubjectHintsEyes, mPerSubjectParentSpeech, criterion = "waic"), rownames = "model")
write.csv(mPerSubjectComparisonWeights, "../saves/supplements-webcamcoding-mPerSubjectComparison-weights.csv", quote = F, row.names = F)
saveRDS(mPerSubjectComparisonWeights, "../saves/supplements-webcamcoding-mPerSubjectComparison-weights.rds")
```

```{r webcam_perTrial_plot}
postSamplesPerSubject <- bind_rows(
  # posterior_samples(mPerTrialNull, pars = "age") %>% mutate(model = "mPerTrialNull - age") %>% rename(estimate = b_age),
  posterior_samples(mPerSubjectPI, pars = "pi") %>% mutate(model = "mPerSubjectPI") %>% rename(estimate = b_pi), 
  posterior_samples(mPerSubjectRepeatedTestquestion, pars = "repeatedTestquestion") %>% mutate(model = "mPerSubjectRepeatedTestquestion") %>% rename(estimate = b_repeatedTestquestion), 
  posterior_samples(mPerSubjectVerificationChoice, pars = "verificationChoice") %>% mutate(model = "mPerSubjectVerificationChoice") %>% rename(estimate = b_verificationChoice), 
  posterior_samples(mPerSubjectHintsEyes, pars = "hintsEyes") %>% mutate(model = "mPerSubjectHintsEyes") %>% rename(estimate = b_hintsEyes), 
  posterior_samples(mPerSubjectParentSpeech, pars = "parentSpeech") %>% mutate(model = "mPerSubjectParentSpeech") %>% rename(estimate = b_parentSpeech), 
)

postSamplesPerSubject <- postSamplesPerSubject %>% 
  mutate(
        model = factor(model, levels = c("mPerSubjectParentSpeech", "mPerSubjectHintsEyes", "mPerSubjectVerificationChoice", "mPerSubjectRepeatedTestquestion", "mPerSubjectPI")),
  )

ggplot(postSamplesPerSubject, aes(x = estimate, y = model, fill = factor(stat(quantile)))) +
  geom_vline(xintercept = 0, lty = 2, alpha = .75) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
  ) +	
  scale_fill_manual(values = c("firebrick", "NA", "firebrick"), guide = "none")

ggsave("../figures/supplements-webcamcoding-mPerSubjectComparison-posteriorsample.png")
```

```{r overall_save}
overall_resuls <- bind_rows(mPerTrialComparison, mPerSubjectComparison) %>% 
  left_join(bind_rows(mPerTrialComparisonWeights, mPerSubjectComparisonWeights)) %>% 
  select(model, waic, value, elpd_diff, se_diff) %>% 
  rename(waic_weight = value)

write.csv(overall_resuls, "../saves/supplements-webcamcoding-resuls.csv", quote = F, row.names = F)
saveRDS(overall_resuls, "../saves/supplements-webcamcoding-results.rds")
```