### 2023-03-10 stats/supplements_triallevel.Rmd

---
title: "TANGO training"
author: "Julia Prein"
date: "02/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(ggthemes)
library(ggrain)

options(scipen = 999)
theme_set(theme_classic())
```

```{r load_data}
alltrials <- readRDS(file = "../data/tango-alltrials.rds") %>% 
  # determine order of factors
  mutate(
    targetPosition = factor(targetPosition, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "box1", "box2", "box3", "box4", "box5")), 
    studyversion = factor(studyversion, levels = c("hedge", "box")), 
    datacollection = factor(datacollection, levels = c("in-person - supervised", "remote - unsupervised")),
    sample = factor(sample, levels = c("kids", "adults")), 
    trialType = factor(trialType, levels = c("touch", "fam", "test"), labels = c("training 1", "training 2 & 3", "test")), 
    studytype = factor(studytype, levels = c("vali", "reli"), labels = c("individual variation sample", "reliability sample")),

    # for models, summarize targetPositions in a way that it's symmetrical for left/ right side
    # the smaller the value, the more central the position
    symmetricPosition = case_when(
                                  targetPosition == "5" | targetPosition == "6" | targetPosition == "box3" ~ 1, 
                                  targetPosition == "4" | targetPosition == "7" | targetPosition == "box2" | targetPosition == "box4" ~ 2,
                                  targetPosition == "3" | targetPosition == "8" | targetPosition == "box1" | targetPosition == "box5" ~ 3, 
                                  targetPosition == "2" | targetPosition == "9" ~ 4,
                                  targetPosition == "1" | targetPosition == "10" ~ 5, 
                       ),

    # distance between target and click in balloon widths (balloon width is 160 SVG units)
    clickDistInBalloons = abs(clickDistFromTargetCenterX)/160,
  )
```

```{r trainingtrials}
alltrials %>% 
  filter(ageInYears < 6) %>% 
  ggplot(., aes(x = trialType, y = clickDistInBalloons, fill = trialType)) +
  geom_rain(cov = "clickDistInBalloons", 
            point.args.pos = rlang::list2(size = 0.5, position = position_jitter(width = 0.05, height = 0)),
            boxplot.args.pos = list(
              width = 0.1, position = position_nudge(x = 0.13)),
            violin.args.pos = list(
              side = "r",
              width = 1, position = position_nudge(x = 0.2))) +
  guides(fill = 'none', color = 'none') +
  scale_colour_gradient(high = "#220220", low = "#D3D3D3") + 
  scale_fill_manual(values=c("#c6d325", "#00b1ea", "#29485d")) + 
  facet_grid(studytype ~ studyversion) + 
  scale_x_discrete(expand = c(0, 0)) +
  labs(x = "Trial type", y = "Imprecision in target width")

ggsave("../figures/tango_supplements_trainingtrials.png", width = 9, height = 6, scale = 1.2, bg = "white")
```

```{r across_trials}
alltrials %>% 
  filter(trialType == "test" & ageInYears < 6) %>%
  ggplot(mapping = aes(x = trialNr, y = clickDistInBalloons)) +
  geom_point(aes(color = subjID, group = subjID), alpha = 0.2) +
  geom_line(aes(color = subjID, group = subjID), alpha = 0.2) +
  geom_smooth(method = "lm", linetype = "dashed", size = 1.5, color = "black", se = FALSE) +
  scale_x_continuous(breaks = seq(1, 19, 1)) + 
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) + 
  xlab("Trial Number") + 
  ylab("Imprecision in target widths") +
  theme_classic() +
  theme(legend.position = "none") + 
  facet_grid(studytype ~ studyversion)

ggsave("../figures/tango_supplements_acrosstrials.png", width = 8, height = 6, scale = 1, bg = "white")
```