### 2023-05-12 stats/supplements_validity.Rmd
incorporate reviews round 2

---
title: "TANGO Validity"
author: "Julia Prein"
date: "2023-03-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(ggpubr)
library(ggthemes)
library(brms) # modeling
library(tidyverse) # data handling etc.
library(tidybayes) # gather_draws
library(ggtext) # for printing italics

options(scipen = 999)
theme_set(theme_classic())
```

```{r load_data}
testtrials <- readRDS(file = "../data/tango-testtrials.rds") %>% 
  # determine order of factors
  mutate(
    targetPosition = factor(targetPosition, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "box1", "box2", "box3", "box4", "box5")), 
    studyversion = factor(studyversion, levels = c("hedge", "box")), 
    datacollection = factor(datacollection, levels = c("in-person - supervised", "remote - unsupervised")),
    sample = factor(sample, levels = c("kids", "adults")), 
    studytype = factor(studytype, levels = c("vali", "vali2", "reli")),

    # for models, summarize targetPositions in a way that it's symmetrical for left/ right side
    # the smaller the value, the more central the position
    symmetricPosition = ifelse(studyversion == "box", 
                               # for box version
                               case_when(
                                  targetPosition == "box3" ~ 0, 
                                  targetPosition == "box2" | targetPosition == "box4" ~ 1,
                                  targetPosition == "box1" | targetPosition == "box5" ~ 2), 
                               # for hedge version
                               targetCentralityX
                       ),

    # distance between target and click in balloon widths (balloon width is 160 SVG units)
    clickDistInBalloons = abs(clickDistFromTargetCenterX)/160,
  )
```

```{r orev_data}
orev_data <- read.table("../data/tango-orev-data.csv", encoding="UTF-8", header = TRUE, sep = ",", quote = "\"", dec = ".") %>% 
  mutate_if(is.character, as.factor)
```

# SOCIAL ENVIRONMENTAL INFLUENCES

```{r socenv_prepareData}
mDataSocenv <- testtrials %>% 
  filter(ageInYears < 6) %>% 
  select(datacollection, studyversion, subjID, ageInMonths, ageInYears, correctBox, childcareAge, childcareHours, householdTotal, siblingsNr, siblingsChildAged, siblingsVarietyScoreP, siblingsVarietyScoreW) %>% 
  group_by(subjID) %>% 
  mutate(siblingsChildAged = replace(siblingsChildAged, siblingsChildAged == -4, 0)) %>% 
  # so that all models run with same nr of subjects
  drop_na() %>%
  ungroup()

socenv_predictordistributions <- mDataSocenv %>% 
  group_by(subjID) %>% 
  slice(1) %>% 
  select(childcareAge, childcareHours, householdTotal, siblingsNr, siblingsChildAged, siblingsVarietyScoreP, siblingsVarietyScoreW) %>% 
  rename(
    "Age of childcare entry in months" = childcareAge, 
    "# Children in household aged 0-18" = siblingsNr, 
    "Sibling variety score (Peterson, 2000)" = siblingsVarietyScoreP, 
    "# Children in household aged 1-12" = siblingsChildAged,
    "# Household members" = householdTotal,
    "Average hours spent in childcare per day" = childcareHours, 
    "Sibling variety score (Cassidy et al., 2005)" = siblingsVarietyScoreW,
  ) %>% 
  psych::describe() %>% 
  select(-c(vars, median, trimmed, mad, range))

socenv_predictordistributions
saveRDS(socenv_predictordistributions, file = "../saves/socenv_predictordistributions.rds")

mDataSocenv <- mDataSocenv %>% 
  mutate(
    # scaled (centered around mean, SD = 1, i.e., z-transformed)
    hhSibs = scale(siblingsNr, center = TRUE, scale = TRUE),
    sibsChildAged = scale(siblingsChildAged, center = TRUE, scale = TRUE),
    ccAge = scale(childcareAge, center = TRUE, scale = TRUE),
    ccHours = scale(childcareHours, center = TRUE, scale = TRUE),
    hhTotal = scale(householdTotal, center = TRUE, scale = TRUE),
    age = scale(ageInMonths, center = TRUE, scale = TRUE), 
    peerExposure = scale(childcareHours, center = TRUE, scale = TRUE) + scale(siblingsNr, center = TRUE, scale = TRUE),
    pe = scale(peerExposure, center = TRUE, scale = TRUE), 
  ) %>% 
  group_by(datacollection, studyversion, subjID, age, ageInMonths, ageInYears, ccAge, ccHours, hhTotal, hhSibs, sibsChildAged, siblingsVarietyScoreP, siblingsVarietyScoreW, pe) %>% 
  summarise(
    cor_tango = sum(correctBox),
    n_tango = n()
  ) %>% 
  mutate_if(is.character, as.factor)
```

```{r socenv_sample}
testtrials %>%
  filter(ageInYears < 6) %>% 
  distinct(subjID, .keep_all = T) %>%
  pivot_longer(cols = c(childcareAge, childcareHours, householdTotal, siblingsNr, siblingsChildAged, siblingsVarietyScoreP, siblingsVarietyScoreW), names_to = "score", values_to = "value") %>%
  group_by(score) %>%
  summarise(na = sum(is.na(value)))

n_distinct(mDataSocenv$subjID)

socenv_sample <- mDataSocenv %>% 
  group_by(ageInYears) %>%
  summarise(
    nTotal = n_distinct(subjID), 
    meanAge = mean(ageInMonths, na.rm = T) %>% round(2),
    sdAge = sd(ageInMonths, na.rm = T) %>% round(2),
    minAge = min(ageInMonths, na.rm = T),
    maxAge = max(ageInMonths, na.rm = T),
  ) %>% 
  mutate(sample = "socenv")

socenv_sample
```

```{r socenv_nullmodel}
mSEnull <- brm(cor_tango | trials(n_tango) ~ age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))
```

```{r socenv_fullmodels}
mSEccAge <- brm(cor_tango | trials(n_tango) ~ ccAge + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mSEccHours <- brm(cor_tango | trials(n_tango) ~ ccHours + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mSEhhTotal <- brm(cor_tango | trials(n_tango) ~ hhTotal + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mSEhhSibs <- brm(cor_tango | trials(n_tango) ~ hhSibs + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mSEsibsChildAged <- brm(cor_tango | trials(n_tango) ~ sibsChildAged + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mSEsiblingsVarietyScoreP <- brm(cor_tango | trials(n_tango) ~ siblingsVarietyScoreP + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mSEsiblingsVarietyScoreW <- brm(cor_tango | trials(n_tango) ~ siblingsVarietyScoreW + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mSEPE <- brm(cor_tango | trials(n_tango) ~ pe + age + datacollection + studyversion,
                data = mDataSocenv,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))
```

```{r socenv_modelcomparisons}
mSEComparison <- loo_compare(mSEnull, mSEccAge, mSEccHours, mSEhhTotal, mSEhhSibs, mSEsibsChildAged, mSEsiblingsVarietyScoreP, mSEsiblingsVarietyScoreW, mSEPE, criterion = "waic")%>% 
  as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mSEnull, mSEccAge, mSEccHours, mSEhhTotal, mSEhhSibs, mSEsibsChildAged, mSEsiblingsVarietyScoreP, mSEsiblingsVarietyScoreW, mSEPE, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mSEComparison

socenvModelorder <- mSEComparison %>% arrange(modelweight) %>% pull(model)
```

```{r socenv_savemodels}
write.csv(mSEnull, "../saves/mSEnull.csv", quote = F, row.names = F)
saveRDS(mSEnull, "../saves/mSEnull.rds")
write.csv(mSEccAge, "../saves/mSEccAge.csv", quote = F, row.names = F)
saveRDS(mSEccAge, "../saves/mSEccAge.rds")
write.csv(mSEccHours, "../saves/mSEccHours.csv", quote = F, row.names = F)
saveRDS(mSEccHours, "../saves/mSEccHours.rds")
write.csv(mSEhhTotal, "../saves/mSEhhTotal.csv", quote = F, row.names = F)
saveRDS(mSEhhTotal, "../saves/mSEhhTotal.rds")
write.csv(mSEhhSibs, "../saves/mSEhhSibs.csv", quote = F, row.names = F)
saveRDS(mSEhhSibs, "../saves/mSEhhSibs.rds")
write.csv(mSEsibsChildAged, "../saves/mSEsibsChildAged.csv", quote = F, row.names = F)
saveRDS(mSEsibsChildAged, "../saves/mSEsibsChildAged.rds")
write.csv(mSEsiblingsVarietyScoreP, "../saves/mSEsiblingsVarietyScoreP.csv", quote = F, row.names = F)
saveRDS(mSEsiblingsVarietyScoreP, "../saves/mSEsiblingsVarietyScoreP.rds")
write.csv(mSEsiblingsVarietyScoreW, "../saves/mSEsiblingsVarietyScoreW.csv", quote = F, row.names = F)
saveRDS(mSEsiblingsVarietyScoreW, "../saves/mSEsiblingsVarietyScoreW.rds")
write.csv(mSEPE, "../saves/mSEPE.csv", quote = F, row.names = F)
saveRDS(mSEPE, "../saves/mSEPE.rds")
```

```{r socenv_save}
socenv_results <- mSEComparison %>%
  select(model, waic, se_waic, modelweight, elpd_diff, se_diff) %>%
  rename(
    Predictor = model, 
    WAIC = waic, 
    SE_WAIC = se_waic,
    Weight = modelweight, 
    ELPD_DIFF = elpd_diff, 
    SE_ELPD_DIFF = se_diff,
  ) %>% 
  mutate(
    Predictor = recode(Predictor, 
                   "mSEPE" = "Peer exposure index", 
                   "mSEsiblingsVarietyScoreW" = "Sibling variety score (Cassidy et al., 2005)", 
                   "mSEhhSibs" = "# Children in household aged 0-18", 
                   "mSEccHours" = "Average hours spent in childcare per day",
                   "mSEnull" = "Null model", 
                   "mSEhhTotal" = "# Household members",
                   "mSEccAge" = "Age of childcare entry", 
                   "mSEsiblingsVarietyScoreP" = "Sibling variety score (Peterson, 2000)", 
                   "mSEsibsChildAged" = "# Children in household aged 1-12",
                   )
  )
write.csv(socenv_results, "../saves/socenv_results.csv", quote = F, row.names = F)
saveRDS(socenv_results, "../saves/socenv_results.rds")
```

```{r socenv_posteriordraws}
socenv_draws <- bind_rows(
  mSEccAge %>% gather_draws(b_ccAge),
  mSEccHours %>% gather_draws(b_ccHours), 
  mSEhhTotal %>% gather_draws(b_hhTotal), 
  mSEhhSibs %>% gather_draws(b_hhSibs), 
  mSEsibsChildAged %>% gather_draws(b_sibsChildAged), 
  mSEsiblingsVarietyScoreP %>% gather_draws(b_siblingsVarietyScoreP), 
  mSEsiblingsVarietyScoreW %>% gather_draws(b_siblingsVarietyScoreW),
  mSEPE %>% gather_draws(b_pe)
  ) %>% 
  mutate(
    predictor = factor(.variable, 
                       levels = c("b_hhSibs", 
                                  "b_siblingsVarietyScoreP", 
                                  "b_sibsChildAged", 
                                  "b_hhTotal", 
                                  "b_ccHours", 
                                  "b_ccAge", 
                                  "b_siblingsVarietyScoreW", 
                                  "b_pe"), 
                       labels = c("# Children in household aged 0-18", 
                                  "Sibling variety score (Peterson, 2000)", 
                                  "# Children in household aged 1-12", 
                                  "# Household members", 
                                  "Average hours spent in childcare per day", 
                                  "Age of childcare entry", 
                                  "Sibling variety score (Cassidy et al., 2005)", 
                                  "Peer exposure index")
                       )  
  ) %>% 
  rename(estimate = .value)

write.csv(socenv_draws, "../saves/socenv_draws.csv", quote = F, row.names = F)
saveRDS(socenv_draws, "../saves/socenv_draws.rds")
```

```{r socenv_plot_winning}
socenv_plot_winning <- socenv_draws %>% 
  filter(predictor == "Age of childcare entry") %>% 
  ggplot(., aes(y = predictor, x = estimate)) +
  tidybayes::stat_halfeye(.width = c(.95, .8), scale = 0.8, slab_color = "#006c66", slab_fill = "#EBECEC", slab_alpha = 0.9) +
  labs(x = "Posterior estimate when predicting TANGO score", y = "") +
  guides(fill = "none")+
  geom_vline(xintercept = 0, linetype = "dashed") +
  coord_cartesian(expand = TRUE, ylim = c(1.5, 1.5)) +
  theme_classic()

socenv_plot_winning

saveRDS(socenv_plot_winning, file = "../saves/socenv_plot_winning.rds")
# ggsave("../figures/socenv_plot_winning.png", width = 10, height = 6, scale = 1, bg = "white")
```

```{r socenv_plot_allother}
socenv_plot_allother <- socenv_draws %>% 
  filter(predictor != "Age of childcare entry") %>% 
  ggplot(., aes(y = predictor, x = estimate)) +
  tidybayes::stat_halfeye(.width = c(.95, .8), scale = 1, slab_color = "#006c66", slab_size = 0.5, slab_fill = "#EBECEC", slab_alpha = 0.9) +
  labs(x = "Posterior estimate when predicting TANGO score", y = "") +
  guides(fill = "none")+
  geom_vline(xintercept = 0, linetype = "dashed") +
  coord_cartesian(expand = TRUE, ylim = c(1.5, 7.5)) +
  theme_classic()

socenv_plot_allother

saveRDS(socenv_plot_allother, file = "../saves/socenv_plot_allother.rds")
# ggsave("../figures/socenv_plot_allother.png", width = 10, height = 6, scale = 1, bg = "white")
```

# RECEPTIVE VOCABULARY

```{r orev_sample}
orev_sample <- orev_data %>% 
  group_by(ageInYears) %>%
  summarise(
    nTotal = n_distinct(subjID), 
    meanAge = mean(age_orev*12, na.rm = T) %>% round(2),
    sdAge = sd(age_orev*12, na.rm = T) %>% round(2),
    minAge = min(age_orev*12, na.rm = T) %>% round(),
    maxAge = max(age_orev*12, na.rm = T) %>% round(),
  ) %>% 
  mutate(sample = "orev")

time_diff <- orev_data %>% 
  summarise(
    meanTimeDiff = mean(time_diff, na.rm = T) %>% round(2),
    sdTimeDiff = sd(time_diff, na.rm = T) %>% round(2),
    minTimeDiff = min(time_diff, na.rm = T) %>% round(2),
    maxTimeDiff = max(time_diff, na.rm = T) %>% round(2),
  ) %>% 
  mutate(sample = "orev_tango")

vali_sample <- bind_rows(socenv_sample, orev_sample, time_diff)

vali_sample
saveRDS(vali_sample, file = "../saves/vali_sample.rds")
```

```{r orev_prepareData}
mDataOrev <- orev_data %>% 
  mutate(
    age_centered = scale(age_orev, scale = T, center = T), 
    mean_tango_centered = scale(mean_tango, scale = T, center = T),
  )

n_distinct(mDataOrev$subjID)
```

```{r orev_models}
mOrevNull <- brm(cor_orev | trials(n_orev) ~ 1,
                 data = mDataOrev,
                 family = binomial, 
                 warmup = 1000, 
                 iter   = 3000, 
                 chains = 4, 
                 inits  = "random",
                 cores  = 4
     ) %>%
  add_criterion(c("loo","waic"))

mOrevAge <- brm(cor_orev | trials(n_orev) ~ age_centered,
                data = mDataOrev,
                family = binomial, 
                warmup = 1000, 
                iter   = 3000, 
                chains = 4, 
                inits  = "random",
                cores  = 4
            ) %>% 
  add_criterion(c("loo","waic"))

mOrevFull <- brm(cor_orev | trials(n_orev) ~ age_centered + mean_tango_centered,
                 data = mDataOrev,
                 family = binomial, 
                 warmup = 1000, 
                 iter   = 3000, 
                 chains = 4, 
                 inits  = "random",
                 cores  = 4
              ) %>%
  add_criterion(c("loo","waic"))
```

```{r orev_modelcomparison}
mOrevComparison <- loo_compare(mOrevNull, mOrevAge, mOrevFull, criterion = "waic") %>% as_tibble(rownames = "model") %>%
  left_join(as_tibble(model_weights(mOrevNull, mOrevAge, mOrevFull, weights = "waic"), rownames = "model")) %>% 
  rename(modelweight = value)

mOrevComparison
```

```{r orev_save}
orev_results <- mOrevComparison %>%
  select(model, waic, se_waic, modelweight, elpd_diff, se_diff) %>%
  rename(
    Predictor = model, 
    WAIC = waic, 
    SE_WAIC = se_waic,
    Weight = modelweight, 
    ELPD_DIFF = elpd_diff, 
    SE_ELPD_DIFF = se_diff,
  ) %>% 
  mutate(
    Predictor = recode(Predictor, 
                   "mOrevFull" = "Age (scaled), TANGO score (scaled)", 
                   "mOrevAge" = "Age (scaled)", 
                   "mOrevNull" = "Null model", 
                   )
  )

write.csv(orev_results, "../saves/orev_results.csv", quote = F, row.names = F)
saveRDS(orev_results, "../saves/orev_results.rds")
```

```{r orev_posteriordraws}
orev_draws <- bind_rows(
  mOrevAge %>% gather_draws(b_age_centered) %>% mutate(model = "age model"), 
  mOrevFull %>% gather_draws(b_age_centered, b_mean_tango_centered) %>% mutate(model = "full model"), 
  ) %>% 
  mutate(
    .variable = factor(.variable, levels = c("b_age_centered", "b_mean_tango_centered"), labels = c("Age (scaled)", "TANGO score (scaled)")), 
  ) %>% 
  rename(estimate = .value, predictor = .variable)

write.csv(orev_draws, "../saves/orev_draws.csv", quote = F, row.names = F)
saveRDS(orev_draws, "../saves/orev_draws.rds")
```

```{r orev_modelplot}
orev_modelplot <- orev_draws %>% 
  ggplot(., aes(x = estimate, y = predictor)) +
 tidybayes::stat_halfeye(.width = c(.95, .8), scale = 1, slab_color = "#006c66", slab_fill = "#EBECEC", slab_alpha = 0.9) +
    labs(x = "Posterior estimate when predicting oREV score", y = "") +
  guides(fill = "none")+
  geom_vline(xintercept = 0, linetype = "dashed") +
  coord_cartesian(expand = TRUE, xlim = c(-0.1, 0.6), ylim = c(1.5, 2.5)) +
  theme_classic()

orev_modelplot

saveRDS(orev_modelplot, file = "../saves/orev_modelplot.rds")
# ggsave("../figures/tango_orev_model.png", width = 10, height = 6, scale = 0.7, bg = "white")
```

```{r orev_plot}
orev_data %>%
  group_by(ageInYears) %>%
  summarize(cor=cor(mean_tango, mean_orev) %>% round(2))

orev_plot <- ggplot(orev_data, aes( x = mean_tango, y = mean_orev)) +

  geom_jitter(aes(col = as.factor(ageInYears)), shape = 18, size = 2, width = 0.05, height = 0.05) +
  geom_smooth(method = "glm", size = 1, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0.9, label.y = 0.98, aes(label = paste(..r.label..)), size = 4, r.accuracy = 0.01, cor.coef.name = "r") +

  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d", "#9D3BBD", "#F71C8D"), labels = c("3-year-olds (*r* = 0.26)", "4-year-olds (*r* = 0.22)", "5-year-olds (*r* = 0.29)", "6-year-olds (*r* = 0.11)", "7-year-olds (*r* = 0.40)")) +

  labs(x = "TANGO score", y = "oREV score", col = "Age in years") + 
  scale_x_continuous(limits = c(0, 1.05), breaks = seq(0, 1, 0.25)) + 
  scale_y_continuous(limits = c(0, 1.1), breaks = seq(0, 1, 0.25)) +

  theme(legend.title = element_text(size = 9), 
        legend.key.size = unit(0.4, "cm"), 
        legend.direction = "vertical", 
        legend.box = "horizontal",
        legend.position = c(0.2, 0.18), 
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"), 
        legend.text = element_markdown()
        )

orev_plot

saveRDS(orev_plot, file = "../saves/orev_plot.rds")
# ggsave("../figures/tango_orev_plot.png", width = 7, height = 4)
```

# ARRANGE PLOT

```{r arrange_plots}
vali_arrangedplot <- ggarrange(
  ggarrange(
    socenv_plot_winning,
    socenv_plot_allother,
    ncol = 2,
    labels = c("A", "B")
  ),

  ggarrange(
    orev_modelplot, 
    orev_plot,
    ncol = 2,
    labels = c("C", "D")
  ), 
  nrow = 2
  )

vali_arrangedplot

saveRDS(vali_arrangedplot, file = "../saves/vali-arrangedplot.rds")
ggsave("../figures/tango_vali_arrangedplot.png", width = 8.5, height = 6, scale = 1.5, bg = "white")
```

```{r orev_by_agegroup}
orev_agegroup <- ggplot(orev_data, aes( x = mean_tango, y = mean_orev)) +

  geom_jitter(aes(col = as.factor(ageInYears)), shape = 18, size = 2, width = 0.05, height = 0.05) +
  geom_smooth(method = "glm", size = 1, alpha = 0.15, col = "darkgrey") +
  stat_cor(method = "pearson", label.x = 0.05, label.y = 1.1, aes(label = paste(..r.label..)), size = 3, r.accuracy = 0.01, cor.coef.name = "r") +

  scale_color_manual(values = c("#c6d325", "#00b1ea", "#29485d", "#9D3BBD", "#F71C8D"), labels = c("3-year-olds (*r* = 0.26)", "4-year-olds (*r* = 0.22)", "5-year-olds (*r* = 0.29)", "6-year-olds (*r* = 0.11)", "7-year-olds (*r* = 0.40)")) +

  labs(x = "TANGO score", y = "oREV score", col = "Age in years") + 
  scale_x_continuous(limits = c(0, 1.05), breaks = seq(0, 1, 0.5)) + 
  scale_y_continuous(limits = c(0, 1.1), breaks = seq(0, 1, 0.25)) +
  facet_grid(~ ageInYears) + 
  theme(legend.position = "none")
d
orev_agegroup

saveRDS(orev_agegroup, file = "../saves/tango_orev_byagegroup.rds")
ggsave("../figures/tango_supplements_orev_byagegroup.png", width = 7, height = 4)
```